Respiratory Disease Outbreaks in Toronto Pre, During, and Post Covid-19 Pandemic 

### 2025-05-13 paper/paper.qmd
previewed rough rough draft

```{python}
#| label: fig-polar-1
#| fig-cap: "A line plot on a polar axis"

@@

For every record there is an id number, institution name, institution address, outbreak setting (type of facility), type of outbreak, causative agent-1, causative agent-2, date outbreak began, date outbreak ended, and whether the outbreak is currently active.

@@

Causative Agent-1 is listed as either as the specific causative agent (eg Covid-19), or is listed as "Unable to Identify". 
Causative Agent-2 is listed as either the specific secondary causative agent, or is listed as "None".

Below is a preview of this information:

For this analysis, the variables of interest are the yearly number of outbreaks for Influenza (A & B), Parainfluenza, Respiratory Syncytial Virus (RSV), Metapneumovirus, and Rhinovirus over the years of 2019 - 2024 inclusive.

@@

For each year, the number of cases were aggregated, and were aggregated within each respective category. The diseases of interest were: Coronavirus Influenza (A & B), Parainfluenza, Respiratory Syncytial Virus (RSV), Metapneumovirus, and Rhinovirus. 
Instances of coronavirus and covid-19 were grouped together under the title of “coronavirus”. 
Influenza A & B were additionally grouped together under the title “Influenza”.

@@

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

And the summary statistics for the cleaned data:

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

@@

## Total Outbreak, Coronavirus Outbreak, and Respiratory Outbreak 
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

## Influenza Outbreak & Other

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

## Other outbreaks

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults-2
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

@@

## Total Outbreak, Coronavirus Outbreak, and Respiratory Outbreak {#sec-first-point}

@@

## Influenza Outbreak & Other

@@

## Other outbreaks

@@

One main weakness with this analysis is it is limited to healthcare facilities, which fail to be representative of the larger demographic of Toronto. A next step would be to analyze disease outbreak in the general population.

@@

>>>>

@@

For every record there is an id number, institution name, institution address, outbreak setting (type of facility), type of outbreak, causative agent-1, causative agent-2, date outbreak began, date outbreak ended, and whether the outbreak is currently active. Below is a preview of this information:

@@

Causative Agent-1 is either a specific agent (ie: Coronavirus), or "Unable to identify".
Causative Agent-2 is either a specific agent, or "None".

*insert table showing first few lines of raw data

For this analysis, the variables of interest are the yearly number of outbreaks for Influenza (A & B), Parainfluenza, Respiratory Syncytial Virus (RSV), Metapneumovirus, and Rhinovirus. Over 2019-2024.

@@

For each year, the number of cases were aggregated, and were aggregated within each respective category. The diseases of interest were: Coronavirus Influenza (A & B), Parainfluenza, Respiratory Syncytial Virus (RSV), Metapneumovirus, and Rhinovirus. 
Instances of coronavirus and covid-19 were grouped together under the title of “coronavirus”, given covid is caused by a coronavirus. Influenza A & B were additionally grouped together under the title “Influenza”.

@@

*** insert table

Below are the summary statistics for the cleaned data:

*** insert summary statistics

@@

Total & respiratory & Coronavirus:
*** insert graph

We see that the total number of cases fluctuates with the number of respiratory cases, with number of outbreaks increasing since 2019 to 2022, and beginning to decrease afterwards

The number of cases also fluctuates with the number of respiratory cases and total cases, following the same trend described above.

Composition of Respiratory Illness:
*** insert graph 
Covid / Secondary Disease / Other & Unknown

Influenza:
Number of cases fall from 2019- 2021, with 0 cases in 2021, and increase thereafter.

Parainfluenza, RSV, Metapneumovirus, and Rhinovirus:
Each has specifics, see general trend of lower number & some w/ slight decrease reported in pre-pandemic & pandemic, with a subsequent increase post.

@@

## Rise in Non-Covid Respiratory Outbreaks {#sec-first-point}

@@

## Decrease in Covid Percentage of Respiratory Outbreaks

@@

@@

Weaknesses and next steps should also be included.

@@

\appendix
